
---
output:
  slidy_presentation: default
  html_document: default
  ioslides_presentation: default
  beamer_presentation: default
---

## Econ 184b, Econometrics, Fall 2021	
### Module 1: Statistics Review

## Population and data: repeated sampling
Example 1: an abstract population

* Population: a Bernoulli distribution with mean $p$ = 0.5
* Data: a random sample of 54 observations from the above population/distribution
* How to think about the population and data? 

## 
* The above population is the distribution of a fair coin. A fair coin gives a random outcome of 1 (head) or 0 (tail) with equal probability.
* However, this does NOT mean that if you flip a fair coin 54 times, you will always get 27 heads and 27 tails! When you flip the coin 54 times, what you get is data: 54 binary numbers. 

Try this in R: 
```{r,results=F}
a = rbinom(54, 1, 0.5)
print(a)
mean(a)
# The rbinom function generates random numbers with the binomial distribution.
```


## 
* In other words: data is a random realization from the population.
  - Goal of econometrics: learn the population from random data.
* From the data, we determine whether the data is from a fair coin.

Try this in R: 
```{r,results=F}
a = rbinom(54, 1, 0.5)
b = rbinom(54, 1, 0.4)
mean(a)
mean(b)
```

## Example 2: a concrete population
Population: country of origin of all students at Brandeis (5771 students)

* The population is 5771 numbers (either 1 "domestic" or 0 "international") and 4161 of these numbers are 1.
* The population is represented by a Bernoulli distribution 
  - mean $p$ = $\frac{4161}/{5771} = 0.7210189
* Data: a random sample of 54 observations from the above population.

##
How to think about the population and data:

* Again, data is a random realization from the population.
* Goal of econometrics: learn the population from random data.
* Of course, if we are an administrator who observes the entire population, then there’s no need to do econometrics.

**Econometrics/statistics is only useful when we do not observe the population.**

## In many cases, we **never** observe the entire population

Example: population unemployment rate in the U.S.

* Goal: find out exactly what percentage of the people are unemployed.
* Population: employment status of everyone in the U.S. (over 300 million)
* Can we observe the population?
  - If we could send a drone to spy on every person in the U.S. and find out their employment status, then yes. But in practice, this cannot possibly be done.

##
* Data: interview 500 people and find out their employment status.
* How to think about the population and the data:
  - Population: 300 million numbers (either 1 “unemployed” or 0
“employed”).
  - This is represented by a Bernoulli distribution with mean p, where p is the percentage of 1 in these 300 million numbers.
  - Data: 500 random realization from the above Bernoulli distribution
