---
title: "In class exercise 1"
output: html_document
---
You are given the following distribution: Possible values of X are 7, 77 and 777 with the following probability:

$P(X=7)=0.7$    
$P(X=77)=0.2$    
$P(X=777)=0.1$  

Please complete the following:

1. Generate a random sample of N=1000 from above that distribution. What is the mean of your sample?

Answer: the sample mean is user-specific. The population mean is 98, and the sample mean is usually not too far away from it. 
```{r,echo=T,eval=T,fig.keep="high",warning=F,fig.height=4.5}
library(ggplot2)
library(ggthemes)
X = sample(c(7,77,777),size=1000,replace=T,prob=c(0.7,0.2,0.1))

mean(X)
```

2. Plot your sample.
```{r,fig.keep="high"}
ggplot(data=as.data.frame(X)) + geom_histogram(aes(x=X)) + theme_wsj()
```

Answer: See plot below. Plot varies by sample.

3. What is the variance of the distribution?

Answer: The variance is 51589. 

4. What is the variance of your sample?

Answer: the sample variance is user-specific. The population variance is 51589, and the sample variance is usually not too far away from it. 
