---
title: "In class exercise 3 Answer Key"
output: html_document
---
Use 5,000 samples, each with 1,000 observations, Verify that the 95% confidence interval calculated from the (7,77,777) distribution covers the true population mean approximately 95% of the time.

### Answer:
First, we create 5,000 samples.
```{r,message=F,warning=F}
library(tidyverse)
library(magrittr)
####
# create 5000 samples
# store the sample mean and variance
####
N = 5000 # number of repetitions
n=1000 # number of observations in each repitition

mean_store = vector(length=N)
sd_store = vector(length=N)
for (i in 1:N) {
  x = sample(c(7,77,777),size=n,replace=T,prob=c(0.7,0.2,0.1))
  mean_store[i] = mean(x)
  sd_store[i] = sd(x)
}
```

Probability that the confidence interval contains the true population mean, $\mu=98$, is given as:
```{r,echo=T}
ci = data.frame(mean_store,sd_store)
mu = 98
ci %<>% mutate(lower_bound = mean_store - 1.96*sd_store/sqrt(n),
               upper_bound = mean_store + 1.96*sd_store/sqrt(n),
               within_CI = ifelse(mu >= lower_bound & mu <= upper_bound,T,F))
sum(ci$within_CI)/N
```

The empirical probability is `r sum(ci$within_CI)/N` - should be close to 95%. 

(Note: the actual sampling distribution is a student's t distribution, so the critical range for a "true" 95% confidence interval should be qt(0.975,n-1) = `r round(qt(0.975,n-1),3)` instead of qnorm(0.975)=`r round(qnorm(0.975),3)`. This is why in most runs the empirical probability is less than 95%.)

